---
title: "Cacao DataProcessing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# 1. Environment
```{r environ, echo=TRUE, message=FALSE, warning=FALSE}

library(timeSeries)
library(reshape2)
library(cluster)
library(gplots)
library(ggfortify)
library(RColorBrewer)
library(ggplot2)
library(tidyverse)
library(base)
library(graphics)
library(stats)
library(fBasics)
library(pvclust)
library(pca3d)
library(randomForest)
library(varSelRF)
library(e1071)
library(VennDiagram)
library(googlesheets4)
library(MetabolAnalyze)
```

#2. Call the data
```{r}
raw.data <- read.csv("Data/Buenaventura05222020.csv")
min.fill <- read.csv("Data/cleanData.csv") %>% 
  select(-1)
```


# Data processing
# 3.3. Data distribution
The density graph allow us to identify if the data present a normal distribution
```{r}
#1. change the variable names 

metaData <- raw.data[, c(1:3)]

cleanData <- cbind(metaData, min.fill)
clData <- cleanData

colnames(cleanData)[4:4402] <- seq(1,4399)

#2. Make a frequency function
clean.melt <- melt(cleanData, id.vars= seq(1, 3), measure.vars= seq(4, 4402)) 

```

```{r}
#3. Extract signal ID
metab.ID <- as.data.frame(colnames(min.fill))
colnames(metab.ID) <- "signal"        

#4. signal name acquisition
clean.melt$signal <- rep(metab.ID$signal, each=54)
head(clean.melt, 56)

```

### Density graph
Data W/O log transformation or scale
```{r density, warning= FALSE}
ggplot(clean.melt, aes(x = value, fill = Time)) + geom_density(alpha = 0.7) + facet_grid(Region ~ ., scales = "free_y") + xlim(-20, 150)
```


## Log transformation
```{r}
clean.melt <- clean.melt %>% 
  mutate(Log_10 = log10(value),
         Log_2 = log2(value),
         Log_n = log(value))
head(clean.melt)
```


## Log10 Density Graph
```{r}
ggplot(clean.melt, aes(x = Log_10, fill = Time)) + geom_density(alpha = 0.7) + facet_grid(Region ~ ., scales = "free_y") + xlim(-10, 10)

```


## Log2 Density Graph
```{r}
ggplot(clean.melt, aes(x = Log_2, fill = Time)) + geom_density(alpha = 0.7) + facet_grid(Region ~ ., scales = "free_y") + xlim(-30, 30)

```

## Natural Log Density Graph
```{r}
ggplot(clean.melt, aes(x = Log_n, fill = Time)) + geom_density(alpha = 0.7) + facet_grid(Region ~ ., scales = "free_y") + xlim(-30, 30)
```

## 2. Data Transform and Scale 

```{r}
# Transform the matrix data to Log 10, Log 2, y log n
data.Log10 <- clData %>% 
  select(-c(1:3)) %>% 
  map_df(log10)

data.Log2 <- clData %>% 
  select(-c(1:3)) %>% 
  map_df(log2)

data.Logn <- clData %>% 
  select(-c(1:3)) %>% 
  map_df(log)
```


# Function for centering and scaling

```{r}
center_colmeans <- function(x) {
  xcenter = colMeans(x)
  x - rep(xcenter, rep.int(nrow(x), ncol(x)))
}


# Autoscale
center_colstd.dev <- function(x) {
  xcenter = colStdevs(x)
  x / rep(xcenter, rep.int(nrow(x), ncol(x)))
}


# Pareto
center_colpareto <- function(x) {
  xcenter = sqrt(colStdevs(x)) 
  x / rep(xcenter, rep.int(nrow(x), ncol(x)))
}

```



# Autoscale Log10

```{r}
df.center <- center_colmeans(data.Log10)
Autoscale.Log10 <- center_colstd.dev(df.center)
```


1. Density Autoscale Log 10
```{r}

A.10 <- cbind(metaData, Autoscale.Log10)
colnames(A.10)[4:4402] <- seq(1,4399)


A10.melt <- melt(A.10, id.vars= seq(1, 3), measure.vars= seq(4, 4402)) 


A10.melt$signal <- rep(metab.ID$signal, each=54)
head(A10.melt, 56)

ggplot(A10.melt, aes(x = value, fill = Time)) + geom_density(alpha = 0.7) + facet_grid(Region ~ ., scales = "free_y") + xlim(-10, 10)

```


# Pareto Scale Log10

```{r}
df.center <- center_colmeans(data.Log10)
Pareto.Log10 <- center_colpareto(df.center)
```


2. Density Pareto Log 10
```{r}

P.10 <- cbind(metaData, Pareto.Log10)
colnames(P.10)[4:4402] <- seq(1,4399)


P10.melt <- melt(P.10, id.vars= seq(1, 3), measure.vars= seq(4, 4402)) 


P10.melt$signal <- rep(metab.ID$signal, each=54)
head(P10.melt, 56)

ggplot(P10.melt, aes(x = value, fill = Time)) + geom_density(alpha = 0.7) + facet_grid(Region ~ ., scales = "free_y") + xlim(-10, 10)

```

# Autoscale Log2

```{r}
df.center <- center_colmeans(data.Log2)
Autoscale.Log2 <- center_colstd.dev(df.center)
```


1. Density Autoscale Log 2
```{r}

A.2 <- cbind(metaData, Autoscale.Log2)
colnames(A.2)[4:4402] <- seq(1,4399)


A2.melt <- melt(A.2, id.vars= seq(1, 3), measure.vars= seq(4, 4402)) 


A2.melt$signal <- rep(metab.ID$signal, each=54)
head(A2.melt, 56)

ggplot(A2.melt, aes(x = value, fill = Time)) + geom_density(alpha = 0.7) + facet_grid(Region ~ ., scales = "free_y") + xlim(-10, 10)

```


# Pareto Scale Log2

```{r}
df.center <- center_colmeans(data.Log2)
Pareto.Log2 <- center_colpareto(df.center)
```


2. Density Pareto Log 2
```{r}

P.2 <- cbind(metaData, Pareto.Log2)
colnames(P.2)[4:4402] <- seq(1,4399)


P2.melt <- melt(P.2, id.vars= seq(1, 3), measure.vars= seq(4, 4402)) 


P2.melt$signal <- rep(metab.ID$signal, each=54)
head(P2.melt, 56)

ggplot(P2.melt, aes(x = value, fill = Time)) + geom_density(alpha = 0.7) + facet_grid(Region ~ ., scales = "free_y") + xlim(-20, 10)

```


# Autoscale Logn

```{r}
df.center <- center_colmeans(data.Logn)
Autoscale.Logn <- center_colstd.dev(df.center)
```


1. Density Autoscale Log n
```{r}

A.n <- cbind(metaData, Autoscale.Logn)
colnames(A.n)[4:4402] <- seq(1,4399)


An.melt <- melt(A.n, id.vars= seq(1, 3), measure.vars= seq(4, 4402)) 


An.melt$signal <- rep(metab.ID$signal, each=54)
head(An.melt, 56)

ggplot(An.melt, aes(x = value, fill = Time)) + geom_density(alpha = 0.7) + facet_grid(Region ~ ., scales = "free_y") + xlim(-10, 10)

```


# Pareto Scale Ln

```{r}
df.center <- center_colmeans(data.Logn)
Pareto.Logn <- center_colpareto(df.center)
```


2. Density Pareto Ln
```{r}

P.n <- cbind(metaData, Pareto.Logn)
colnames(P.n)[4:4402] <- seq(1,4399)


Pn.melt <- melt(P.n, id.vars= seq(1, 3), measure.vars= seq(4, 4402)) 


Pn.melt$signal <- rep(metab.ID$signal, each=54)
head(Pn.melt, 56)

ggplot(Pn.melt, aes(x = value, fill = Time)) + geom_density(alpha = 0.7) + facet_grid(Region ~ ., scales = "free_y") + xlim(-15, 10)

```


### write final processed data
Log2 and autoscale data processing was used based on the recent literature. But very mild differences were found with log10 transformation.

```{r}
df.autoscale <- Autoscale.Log2
write.csv(df.autoscale, file = "Data/df_autoscale.csv")
write.csv(metaData, file = "Data/metaData.csv")
write.csv(data.Log2, file = "Data/data_Log2.csv")
write.csv(clData, file = "Data/clData.csv")
```




